---
title: "Functions"
author: "Carey Kopeikin"
date: "4/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```


1. Percentages 

a) Use R to find out what percentage .25 is (this should be very easy).

```{r}
.25 *100
```


b) Create a function that takes one argument, a proportion, and returns the percentage.
```{r}
percent.fun <- function( proportion){
  
  percentage <- proportion * 100
  return(percentage)
}

```

```{r}
percent.fun( .3)
```




2. That Right Triangle Formula

a) If you know that the sides adjacent to the right angle of a triangle are 5 and 12 what is the hypotenuse? Use R to find the answer.

```{r}
sqrt( 5^2 + 12^2)
```



Create a function that takes 2 arguments the sides of sides adjacent to the right angle of a triangle and returns the hypotenuse.

```{r}
triangle <- function( a, b){
  
  c = sqrt( a^2 + b^2 ) 
  return(c)
  
}
```

```{r}
triangle( 5, 12)
```




In the dice rolling problems we saw how useful the function ```sample``` can be. As a reminder ```sample``` can pick a value out at random. 

```{r}
?sample
```


Some examples:

This is the same as rolling a six sided die:
```{r}
sample( c(1,2,3,4,5,6), 1)
```

This is also the same as rolling a six sided die:
```{r}
sample( 1:6, 1)
```

All of these are the same as flipping a coin:
```{r}
sample( c(1,2), 1)
sample( 1:2, 1)
sample( c("Heads", "Tails"), 1)
```

By default ```sample``` has the argument replace = FALSE. This means that if it samples multiple numbers it will not replace them. 

For example if I wanted to roll a 6 sided die 6 times it would give me each number once:
```{r}
sample( 1:6, 4)
```
 
 But if I changed the default to TRUE it would give me 6 rolls of the die. 
```{r}
sample( 1:6, 4, replace = TRUE)
```

 


3. In Powerball there are 69 possible numbers for the white balls and 26 possible results for the Powerball. Thus, the odds of picking that perfect combination with a single ticket are one in 292,201,338. Each Powerball ticket costs $2.

a. Generate the numbers of 5 white balls (note there can't be repeats)
```{r}
sample( 1:69, 5, replace = FALSE)
```

b. Generate the Powerball number
```{r}
sample( 1:26, 1)
```

c. Now create a function that takes 2 arguments: the number of white balls and the number of Powerballs and outputs a potential Powerball ticket.

```{r}
powerball <- function( ){
  
  w.numb <- sample( 1:69, 5, replace = FALSE)
  p.numb <- sample( 1:26, 1, replace = FALSE)
  
  return( c(w.numb, p.numb))

}
```

```{r}
powerball()
```





```{r}
powerball <- function( w.balls, p.balls){
  
  w.numb <- sample( 1:w.balls, 5, replace = FALSE)
  p.numb <- sample( 1:p.balls, 1, replace = FALSE)
  return( c(w.numb, p.numb))

}
```

```{r}
powerball( 100, 15)
```





```{r}
powerball( 69, 26)
powerball( 100,5)
```



If statements 
We also saw how useful if statements are. An if statement can be especially useful in a function. Here is a function that takes an argument that is a number and tells you if it is greater than 10:

```{r}
greater.than.ten <- function( my.number ){
  
  if( my.number > 10 ){ return( "Greater than 10") }
  if( my.number == 10){ return( "Exactly 10")}
  
  return("Less than 10")
  
}
```

```{r}
greater.than.ten(3)
```




Now we can see if it worked:
```{r}
greater.than.ten( 80 )
greater.than.ten(10)
greater.than.ten(-5)
greater.than.ten(100)
```

4. Here we will create a guessing game

a. Make R choose a random number between 1 and 5
```{r}
sample(1:5, 1)

```

b. Create a function that takes a single argument that is a number and returns that number.
```{r}
dumb.fun <- function( x){
  
  return(x)
}
```

c. Create a function that:
-   takes a single argument that is a guessed number
-   then picks a random number between 1 and 5
-   then checks to see if the random number is equal to the guessed number
-   then outputs if the guess was correct.
```{r}

guess.game<- function( guess ){
  
  random.num <- sample(1:5, 1)
  if(random.num == guess){return("correct!")}
  
  return("wrong!")

}

```

```{r}
guess.game(4)
```



5. Slots

a. Have R pick 3 numbers between 1 and 3 with replacement.
```{r}
three.numbers <- sample(1:3, 3, replace = TRUE)
```

```{r}
three.numbers
three.numbers[1]
three.numbers[2]
three.numbers[3]
```



b. Create a function that:
-   takes as an argument a number n representing how much you are betting. 
-   picks 3 numbers between 1 and 3
-   checks to see if they are all the same
-   returns whether you won or lost the game of slots
-   if you won tells you that you won 10 times your bet.



```{r}
slots <- function( bet){
  
  numbers <- sample(1:3, 3, replace = TRUE)
  
  if(numbers[1] == numbers[2] & numbers[1] == numbers[3]){ return( c("won", 10*bet) ) }
  
  return("lost")
  
}
```

```{r}
slots(6)
```


```{r}
numbers <- sample(1:3, 3, replace = TRUE)
numbers
numbers[1]
```


```{r}
slots( 15)
```

```{r}
slots <- function( bet){
  
  A <- sample(1:3,1)
  B <- sample(1:3,1)
  C <- sample(1:3,1)
  
  if( A == B & B == C){ return( c("win", 10*bet))}
  
  return("loser!!!")
  
}
```

```{r}
slots( 100 )
```

