---
title: "Day 07 Quiz on Hypothesis Tests"
author: "Carey Kopeikin"
date: "4/3/2022"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```


1. A start-up company is about to market a new computer printer. It decides to gamble by running commercials during the Super Bowl. The company hopes that name recognition will be worth the high cost of the ads. The goal of the company is that over 36% of the public recognize its brand name and associate it with computer equipment. The day after the game, a pollster contacts 420 randomly chosen adults and finds that 181 of them know that this company manufactures printers. Would you recommend that the company continue to advertise during Super Bowls? Explain.

a) Write appropriate hypotheses.
$$H_0: p = .36 \\ H_a: p > .36 $$

b) What is a type I error in this context? Explain the ramifications of that error.

They think over 36% recognize the brand but in reality less than (or equal to) 36% recognizes it. They are happy with the ad and may buy more in the future even though it is a waste of money.


c) What is a type II error in this context? Explain the ramifications of that error.

They think less than (or equal to) 36% recognizes the brand but in reality more than 36% recognizes it.They won't buy more adds even though they are working.


d) Which is worse for the company in this context?

Type I because they will waste a lot of money (other answers could work here).

e) Based on your answer to part d) is it better to use an alpha level of .05 or .01.

.01 because I said Type I was worse and a low alpha results in low Type I error.

f) Perform the test and find the p-value using the alpha you selected in part e.

Define variables
```{r}
# number of successes
x <- 181

# sample size
n <- 420

# sample proportion of successes
phat <- x/n 

# sample proportion of failures
qhat <- 1-phat

# assumed true proportion of successes
p = .36

#assumed true proportion of failures
q = 1-p

#Standard deviation of the sampling distribution of sample proportions
SDphat <- sqrt( (p*q)/n )
```

```{r}
prop.test( x, n, p, alternative = "greater", correct = FALSE)
```
 

g) Make a graph 

```{r}
curve( dnorm( x, p, SDphat),
       xlim=c(p-4*SDphat, p+4*SDphat),
       lwd = 3, 
       main="Sampling Distribution of The \n Sampling Proportion of People who Recognize the Company",
       xlab="Proportion of People who Recognize the Company",
       ylab="Density",
       col = "purple",
       )


segments(phat, 0, phat, dnorm( phat, p, SDphat),
         col="red",
         lwd=5)

polygon(x=c(phat, 
            seq(from=phat, to=1, by=.0001),
            1),
        y=c(0,
            dnorm(seq(from=phat, to=1, by=.0001),mean=p, SDphat),
            0),
        col="yellow")

```
h) State your conclusion.

My p-value of 0.001225 is less than my alpha of .01 so I reject the null hypotheses and accept the alternative that over 36% of the public recognize the company's brand name and associates it with computer printers.



2. For over a year, from the spring of 1721 until winter 1722, a smallpox epidemic afflicted the city of Boston. Though tragic, the 1721 epidemic led to a major milestone in the history of vaccination and smallpox eradication. The use of inoculation during this epidemic, and the heated debate that arose surrounding the practice, was one of the first major applications of inoculations in western society, paving the way for Edward Jenner to develop smallpox vaccination by the end of the century.* 

The file ```smallpox.csv``` is a sample of some of the first people in Boston to be inoculated. Perform a hypothesis test to discover if these first inoculations worked. Note that the mortality rate (death rate) of those who were not vaccinated was 14%.  


a) Write appropriate hypotheses.

$$H_0: p = .14 \\ H_a: p < .14 $$

b) What is a type I error in this context? Explain the ramifications of that error.



c) What is a type II error in this context? Explain the ramifications of that error.


d) Which is worse for the vaccine creator in this context?


e) Based on your answer to part d) is it better to use an alpha level of .05 or .01.


f) Perform the test and find the p-value using the alpha you selected in part e.

```{r}
small <- read.csv("smallpox.csv")
head(small)

```

```{r}
# number of successes
x <- sum(small$result== "died")

# sample size
n <- length(small$result)

# sample proportion of successes
phat <- x/n 

# sample proportion of failures
qhat <- 1-phat

# assumed true proportion of successes
p = .14

#assumed true proportion of failures
q = 1-p

#Standard deviation of the sampling distribution of sample proportions
SDphat <- sqrt( (p*q)/n )
```

```{r}
prop.test( x, n, p, alternative = "less", correct = FALSE)
```

g) Make a graph

```{r}

curve( dnorm( x, p, SDphat),
       xlim=c(p-4*SDphat, p+4*SDphat),
       lwd = 3, 
       main="Sampling Distribution of The \n Sampling Proportion of Should Nots",
       xlab="Proportion of Should Nots",
       ylab="Density",
       col = "purple",
       )


segments(phat, 0, phat, dnorm( phat, p, SDphat),
         col="red",
         lwd=5)

polygon(x=c(0, 
            seq(from=-1, to=phat, by=.0001),
            phat),
        y=c(0,
            dnorm(seq(from=-1, to=phat, by=.0001),mean=p, sd=SDphat),
            0),
        col="yellow")

```


h) State your conclusion.



* https://sitn.hms.harvard.edu/flash/special-edition-on-infectious-disease/2014/the-fight-over-inoculation-during-the-1721-boston-smallpox-epidemic/#:~:text=The%20Boston%20Epidemic,the%20most%20deadly%20%5B2%5D.  




```{r}
small.died <- small[ small$inoculated == "yes",]

# number of successes
x <- sum(small.died$result== "died")

# sample size
n <- length(small.died$result)

# sample proportion of successes
phat <- x/n 

# sample proportion of failures
qhat <- 1-phat

# assumed true proportion of successes
p = .14

#assumed true proportion of failures
q = 1-p

#Standard deviation of the sampling distribution of sample proportions
SDphat <- sqrt( (p*q)/n )
```

```{r}
prop.test( x, n, p, alternative = "less", correct = FALSE)
```
