---
title: "Untitled"
author: "Carey Kopeikin"
date: "4/3/2022"
output: pdf_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```


1. A study in 1994 examined dogs to determine whether there is an increased risk of cancer in dogs that are exposed to the herbicide 2,4-Dichlorophenoxyacetic acid (2,4-D). The results of the dogs who were exposed are in ```cancer_in_dogs1.csv```. In general 32% of dogs who are not exposed to the herbicide develop cancer. 

a) Write appropriate hypotheses.

$$H_0: p = .32 \\ H_a: p >.32$$

b) What is a type I error in this context? Explain the ramifications of that error.
A type I error is if the true p = .32 but we think it is greater than .32. This would mean that we think the herbicide causes cancer even though it doesn't. 


c) What is a type II error in this context? Explain the ramifications of that error.
A type II error is if the true p > .32 but we think that it is = .32. This would mean that we think the herbicide does not cause cancer even though it does.

d) Which is worse in this context?
Up to you

e) Based on your answer to part d) is it better to use an alpha level of .05 or .01.
If you said Type I then .01
If you said Type II then .05

f) Perform the test and find the p-value using the alpha you selected in part e.

```{r}
dogs.data <- read.csv("cancer_in_dogs1.csv")
head(dogs.data) 
# number of successes
x <- sum(dogs.data$response == "cancer")

# sample size
n <- length(dogs.data$response)

# sample proportion of successes
phat <- x/n 

# sample proportion of failures
qhat <- 1-phat

# assumed true proportion of successes
p = .32

#assumed true proportion of failures
q = 1-p

#Standard deviation of the sampling distribution of sample proportions
SDphat <- sqrt( (p*q)/n )
```

```{r}
prop.test(x, n, p, alternative = "greater", correct = FALSE)
```

g) Make a graph

```{r}
curve( dnorm( x, p, SDphat),
       xlim=c(p-4*SDphat, p+4*SDphat),
       lwd = 3, 
       main="Sampling Distribution of The \n Sampling Proportion of Dogs with Cancer",
       xlab="Proportion of Dogs with Cancer",
       ylab="Density",
       col = "purple",
       )


segments(phat, 0, phat, dnorm( phat, p, SDphat),
         col="red",
         lwd=5)

polygon(x=c(phat, 
            seq(from=phat, to=1, by=.0001),
            1),
        y=c(0,
            dnorm(seq(from=phat, to=1, by=.0001),mean=p, SDphat),
            0),
        col="yellow")

```

h) State your conclusion.
Our p value of 0.0008415 is less than our alpha of (whatever you had for part e) so we reject the null hypothesis that the true proportion of dogs with cancer that were exposed to the herbicide 2,4-Dichlorophenoxyacetic acid is .32 and accept the alternative that the true proportion is greater than .32. There is strong evidence that the herbicide causes cancer in dogs.


2. Californians love there In-N-Out Burgers. It is so popular that your friend from California states that 70% of Californians would say In-N-Out makes the best burgers. You think it is less and Collect a survey of 500 respondents. The file ```burger.csv``` contains the results. We have information on their gender and their favorite burger joint. Set up an appropriate hypothesis test to see who wins the bet? 

$$H_0: p = .7 \\ H_a: p <.7$$

b) What is a type I error in this context? Explain the ramifications of that error.
A type I error is if the true p = .7 but we think it is less than .7. This would mean that we would win the bet even though our friend was right.


c) What is a type II error in this context? Explain the ramifications of that error.
A type II error is if the true p < .7 but we think that it is .7. This would mean that our friend would win the bet even though we were right. 

d) Which is worse for you in this context?
Type II is way worse as it means that we would lose the bet even though we were right.


e) Based on your answer to part d) is it better to use an alpha level of .05 or .01.

Type II is worse so .05

f) Perform the test and find the p-value using the alpha you selected in part e.

```{r}
burger.data <- read.csv("burger.csv")
head(burger.data) 
# number of successes
x <- sum(burger.data$best_burger_place == "In-N-Out Burger")

# sample size
n <- length(burger.data$best_burger_place)

# sample proportion of successes
phat <- x/n 

# sample proportion of failures
qhat <- 1-phat

# assumed true proportion of successes
p = .7

#assumed true proportion of failures
q = 1-p

#Standard deviation of the sampling distribution of sample proportions
SDphat <- sqrt( (p*q)/n )
```

```{r}
prop.test(x, n, p, alternative = "less", correct = FALSE)
```

g) Make a graph

```{r}

curve( dnorm( x, p, SDphat),
       xlim=c(p-4*SDphat, p+4*SDphat),
       lwd = 3, 
       main="Sampling Distribution of The \n Sampling Proportion of People who Prefer In-N-Out Burgers",
       xlab="Proportion of People who Prefer In-N-Out Burgers",
       ylab="Density",
       col = "purple",
       )


segments(phat, 0, phat, dnorm( phat, p, SDphat),
         col="red",
         lwd=5)

polygon(x=c(0, 
            seq(from=-1, to=phat, by=.0001),
            phat),
        y=c(0,
            dnorm(seq(from=-1, to=phat, by=.0001),mean=p, sd=SDphat),
            0),
        col="yellow")

```


h) State your conclusion.
Our p value of 0.2473 is greater than our alpha of .05 so we fail to reject the null hypothesis that the true proportion of people who prefer In-N-Out Burgers is equal to .7 and can not say that the true proportion is less than .7 . Technically we could tell our friend that since we failed to reject the null hypothesis rather than accepted it (remember we never accept the null hypothesis) neither of us wins.



