---
title: "Functions"
author: "Carey Kopeikin"
date: "4/09/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```


1. Percentages 

a) Use R to find out what percentage .25 is (this should be very easy).

```{r}

```


b) Create a function that takes one argument, a proportion, and returns the percentage.
```{r}

```




2. That Right Triangle Formula

a) If you know that the sides adjacent to the right angle of a triangle are 5 and 12 what is the hypotenuse? Use R to find the answer.
```{r}

```



Create a function that takes 2 arguments the sides of sides adjacent to the right angle of a triangle and returns the hypotenuse.
```{r}

```



In the dice rolling problems we saw how useful the function ```sample``` can be. As a reminder ```sample``` can pick a value out at random. 

```{r}
?sample
```


Some examples:

This is the same as rolling a six sided die:
```{r}
sample( c(1,2,3,4,5,6), 1)
```

This is also the same as rolling a six sided die:
```{r}
sample( 1:6, 1)
```

All of these are the same as flipping a coin:
```{r}
sample( c(1,2), 1)
sample( 1:2, 1)
sample( c("Heads", "Tails"), 1)
```

By default ```sample``` has the argument replace = FALSE. This means that if it samples multiple numbers it will not replace them. 

For example if I wanted to roll a 6 sided die 6 times it would give me each number once:
```{r}
sample( 1:6, 6)
```
 
 But if I changed the default to TRUE it would give me 6 rolls of the die. 
```{r}
sample( 1:6, 6, replace = TRUE)
```

 


3. In Powerball there are 69 possible numbers for the white balls and 26 possible results for the Powerball. Thus, the odds of picking that perfect combination with a single ticket are one in 292,201,338. Each Powerball ticket costs $2.

a. Generate the numbers of 5 white balls (note there can't be repeats)
```{r}

```

b. Generate the Powerball number
```{r}

```

c. Now create a function that takes 2 arguments: the number of white balls and the number of Powerballs and outputs a potential Powerball ticket.
```{r}

```


If statements 
We also saw how useful if statements are. An if statement can be especially useful in a function. Here is a function that takes an argument that is a number and tells you if it is greater than 10:

```{r}
greater.than.ten <- function( my.number ){
  
  if( my.number > 10 ){ return( "Greater than 10") }
  if( my.number == 10){ return( "Exactly 10")}
  
  return("Less than 10")
  
}
```

Now we can see if it worked:
```{r}
greater.than.ten( 80 )
greater.than.ten(10)
greater.than.ten(-5)
greater.than.ten(100)
```

4. Here we will create a guessing game

a. Make R choose a random number between 1 and 5
```{r}

```

b. Create a function that takes a single argument that is a number and returns that number.
```{r}

```

c. Create a function that:
-   takes a single argument that is a guessed number
-   then picks a random number between 1 and 5
-   then checks to see if the random number is equal to the guessed number
-   then outputs if the guess was correct.
```{r}

```




5. Slots
a. Have R pick 3 numbers between 1 and 3 with replacement.
```{r}

```

b. Create a function that:
-   takes as an argument a number n representing how much you are betting. 
-   picks 3 numbers between 1 and 3
-   checks to see if they are all the same
-   returns whether you won or lost the game of slots
-   if you won tells you that you won 10 times your bet.





