---
title: "ChiSqQuiz.Rmd"
author: "Matthew Bardoe & Carey Kopeikin"
date: "3/19/2020"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

### **Honor Code:** 

This quiz is open note, open book, open internet. You may not seek help from or give help to anyone. You may ask questions of me but if I answer them I am likely to take off points. I will tell you that I will take off points before I answer the question and give you the option of losing the points or trying to figure out the question on your own.  


Type your name below indicating you are signing the honor code.

On my honor I have given nor received unauthorized aid.
*(type name here to sign)*


### Instructions

For each question in this quiz you should use a chi square goodness of
fit test to see if the observed values match the expected values. For
all tests you should list your hypotheses, run the test, explain your
results, and check the residuals if applicable. You should use .05 as
your alpha value for all the tests.

*If you are stuck on any part raise your hand and I may be able to help you at the cost of a few points.*


1.  Sample of students and their GPA and IQ
    <https://www.openintro.org/data/index.php?data=gpa_iq>

A data study was done on students GPA and their IQ. The results are in
the file `gpa_iq.csv` Some question whether it had an appropriate gender
balance. Do a chi square goodness of fit test to see if the gender
balance was representative of the population. You may assume that the
population in this region has 50% males and 50% females. Note that in the data 1
represents male and 2 represents female.



a)  list the hypotheses
H0: the gender of the participants is representative of the population
Ha: the gender of the participants is not representative of the population

b)  run the test
```{r}
study <- read.csv( "gpa_iq.csv")
head(study)
```


```{r}
table(study$gender)
```

```{r}
chisq.test( table(study$gender), p= c(.5, .5) )
```



c)  explain your results

d)  check the residuals and explain the results (if applicable)


2.  NBA Finals
When two competing teams are equally matched, the probability that
each team wins any game is 0.5. The NBA championship goes to the
team that wins four games in a best-of-seven series. If the teams
were equally matched, the probability that the final series ends
with one of the teams sweeping four straight games would be 0.125 .
Further probability calculations indicate that 25% of these series
should last five games, 31.25% should last six games, and the other
31.25% should last the full seven games. The table shows the number
of games it took to decide each of the last 57 NBA champs. Do you
think the teams are usually equally matched? Give statistical
evidence to support your conclusion.    

| Length of series | 4 games |  5 games| 6 games | 7 games |
|------------------|---------|---------|---------|---------|
|  NBA finals      | 7       | 13      | 22      | 15      |


a)  list the hypotheses

H0: the teams were evenly matched
Ha: the teams were not evenly matched

b)  run the test
```{r}
observed <- c( 7, 13, 22, 15)
expectd.percent <- c( .125, .25, .3125, .3125)

chisq.test( observed, p= expectd.percent)
```


c)  explain your results

d)  check the residuals and explain the results (if applicable)


3. Bio Test
A biology professor reports that historically grades in her introductory biology course have
been distributed as follows: 15% A’s, 30% B’s, 40% C’s, 10% D’s, and 5% F’s. Grades in her
most recent course were distributed as follows:

| Grade     | A   | B   | C   | D   | F   |
|-----------|-----|-----|-----|-----|-----|
| Frequency | 89  | 121 | 78  | 25  | 12  |

Test an appropriate hypothesis to decide if the professor’s most recent grade distribution
matches the historical distribution. Give statistical evidence to support your conclusion.

a)  list the hypotheses

H0: the grades are consistent with previous years
Ha: the grades are not consistent with previous years

b)  run the test

```{r}
observed <- c( 89, 121, 78, 25, 12)
expectd.percent <- c( .15, .3, .4, .1, .05)

chi.test.grades <- chisq.test( observed, p= expectd.percent)
chi.test.grades
```


c)  explain your results

d)  check the residuals and explain the results (if applicable)

```{r}
chi.test.grades$residuals
```

