---
title: "Day 16 Functions and Simulations Quiz"
author: "Carey Kopeikin"
date: "1/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```

### **Honor Code:** 

This quiz is open note, open book, open internet. You may not seek help from or give help to anyone. You may ask questions of me but if I answer them I am likely to take off points. I will tell you that I will take off points before I answer the question and give you the option of losing the points or trying to figure out the question on your own.  


Type your name below indicating you are signing the honor code.

On my honor I have given nor received unauthorized aid.
*(type name here to sign)*


*Instructions*



1. *Solitaire*

a. My father's favorite way to waste time is to play solitaire on the computer and he wins 80% of his games. Before going to bed he needs to win 5 games. Create a function that uses a loop to simulate how many games he has to play until he wins 5 games. The function does not need any arguments. The output should be the number of games he needed to play in order to get 5 wins. 


b. On average how many games does he have to play in order to win 5 games. To answer this question run the function a thousand times and report the average games he must play to win 5 games.
```{r}
set.seed(1)
```





4. *Covid*

The following function answers this question:

According to the Washington Post 60% of Americans have had Covid at least once. Create a function that simulates how many students in a class of 20 have had Covid. 


```{r}
covid <- function(  ){
  
  have.covid <- 0
  student.number <- 0
  
  repeat{
    
    student.number <- student.number + 1
    
    covid.test <- sample(1:100, 1)
    
    if( covid.test <= 60){ have.covid <- have.covid +1 }
    
    if( student.number == 20){ break}
  }
  return( have.covid)
}
```




Change the function above so that it answers the following question:

A certain percentage of Americans have had Covid at least once. Create a function that takes two arguments, the percentage of Americans that have had Covid and the number of students in a class. The function should simulate how many of those students have had Covid. 

















