---
title: "Day 15 Functions and Simulations Review"
author: "Carey Kopeikin"
date: "1/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```


1. *Playoff Meltdown*
a) Certain players are unfairly accused of "choking" in the playoffs. A player may just be unlucky during a critical game. Consider Ben Simmons, who did very poorly in the 2020-2021 NBA playoffs. In order to see if his shooting was an aberration create a function that takes two arguments, the number of free throws taken by the player in the playoffs and the player's regular season free throw percentage. The function should then simulate the number of free throws taken assuming the regular season free throw percentage is valid. The function should return the number of made free throws.  

2. *Playoff Meltdown with Loop*

```{r}
bs.loop <- function( attempts, percent){
  
  ft.taken <- 0
  ft.made <- 0
  
  repeat{
    
    ft.taken <- ft.taken + 1
    free.throw <- sample(1:100, 1)
    if(free.throw <= percent){ ft.made <- ft.made + 1}
    
    if(ft.taken == attempts){break}
  }
  
  return(ft.made)
}

```
b) Ben Simmons regular season free throw percentage is 61%. In the 2020-2021 NBA playoffs he took 73 shots. Use your function to simulate how many shots he should have made. Simulate at least 1000 times. 


```{r}
thousand.bs.loop <- replicate( 1000, bs.loop( 73, 61) )
table(thousand.bs.loop)

```

c) In the 2020-2021 NBA playoffs Ben Simmons made only 25 shots. Was he just unlucky or was something else going on?

In 1000 simulations we never got a number close to as small as 25. This suggests that something else was going on.

