---
title: "Functions"
author: "Carey Kopeikin "
date: "4/02/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```


1. Write a function that takes as an input a temperature in Celsius and outputs that temperature in Fahrenheit. 

```{r}
Celsius.to.Fahrenheit <- function( temp.in.Celsius ){
  
  temp.in.Fahrenheit <- temp.in.Celsius * 9/5 + 32
  
  return(temp.in.Fahrenheit)
  
}
```

```{r}
Celsius.to.Fahrenheit(0)
Celsius.to.Fahrenheit(100)
```


2. Write a function that takes a vector as an input and randomly selects three elements from that vector.

```{r}
terrible.team <- read.csv("terrible team.csv")
```


```{r}
three.elements.fun <- function( my.vector ){
  
  elements <- sample( my.vector, 3)
  
  return(elements)
  
}
```

```{r}
three.elements.fun(terrible.team$H)
```

3. Write a function that takes two numbers as inputs and then randomly picks a number in-between those numbers. If no arguments are entered have the default be 5 and 10.

```{r}
in.between.fun <- function( number1=5, number2=10 ){
  
  my.number <- sample( number1:number2, 1)
  
  return(my.number)
  
}
```

```{r}
in.between.fun(30, 100)
in.between.fun()
```


4. Write a function that takes as an input a vector and outputs the number of elements in the vector, the largest number in the vector, and the average of the vector.

```{r}
vector.info <- function( my.vector ){
  
  numb.elements <- length( my.vector)
  largest.num <- max( my.vector)
  average <- mean(my.vector)
  
  return( c( numb.elements, largest.num, average) )
  
}
```

```{r}
vector.info(terrible.team$H)
```


5. Write a function that allows you to enter an integer and outputs the next 4 integers.

```{r}
# method 1
next.four.fun <- function( my.number ){
  
  numbers <- (my.number + 1):(my.number + 4)
  
  return(numbers)
  
}

#method 2

next.four.fun.2 <- function( my.number ){
  
  numbers <- c( (my.number + 1), (my.number + 2), (my.number + 3), (my.number + 4) )
  
  return(numbers)
  
}

#there are many other ways

```

```{r}
next.four.fun( 8)
next.four.fun.2(20)
```

