---
title: "Practice Problems"
author: "Carey Kopeikin"
date: "3/21/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE)
```



## Your Turn

The csv ```get_it_dunn_run.csv``` contains information about the Get it Dunn Run https://www.openintro.org/data/index.php?data=get_it_dunn_run. 


2. The planners of the run expected that 85% of those running would be doing the 5k but in reality it was closer to 79%. Could they just have gotten unlucky or were the numbers significantly different than what was expected?


a) Write appropriate hypotheses.

$$H_0: p = .85 \\ H_a: p<.85    $$

b) Check the assumptions and conditions.
skip

c) Perform the test and find the p-value.
Define variables
```{r}
run <- read.csv("get_it_dunn_run.csv")


# number of successes
x <- sum(run$race == "5k")

# sample size
n <- length(run$race)

# sample proportion of successes
phat <- x/n 

# sample proportion of failures
qhat <- 1-phat

# assumed true proportion of successes
p = .85

#assumed true proportion of failures
q = 1-p

#Standard deviation of the sampling distribution of sample proportions
SDphat <- sqrt( (p*q)/n )
```

```{r}
prop.test( x, n, p=.85, alternative = "less")
```


d) Make a graph
```{r}
#shade to the left

curve( dnorm( x, p, sqrt(p*q/n)),
       xlim=c(p-4*sqrt(p*q/n), p+4*sqrt(p*q/n)),
       lwd = 3, 
       main="Sampling Distribution of The \n Sampling Proportion of 5k runners",
       xlab="Proportion of 5k runners",
       ylab="Density",
       col = "purple",
       )


segments(phat, 0, phat, dnorm( phat, p, sqrt(p*q/n)),
         col="red",
         lwd=5)

polygon(x=c(0, 
            seq(from=-1, to=phat, by=.0001),
            phat),
        y=c(0,
            dnorm(seq(from=-1, to=phat, by=.0001),mean=p, sd=sqrt(p*q/n)),
            0),
        col="yellow")
```

e) State your conclusion.

Since my p-value was less than my alpha of .05 I reject the null hypothesis that the true proportion of runners was .85 and accept the alternative hypothesis that it was less than .05

f) Do you think this difference is meaningful? Explain.
6% seems like a lo to me. I imagine that race organizers may have bought the wrong amount of t-shirts and other memorabilia to sell. 

g) What is a type I error in this context?
When the true proportion is equal to .85 but we think it is less than that. If I was the race coordinator this would mean that I would yell at and possibly fire the person who predicted the amount would be .85 even though they did nothing wrong. 

h) What is a type II error in this context?
When the true proportion is less than .85 but we think it is .85. If I was the race coordinator this would mean that I would continue to trust my incompetent employees. 

i) Which is worse in this context?
This is up to you.

j) Based on your answer to part i) is it better to use an alpha level of .1 or .001.
If you said Type I is worse .001
If you said Type II is worse .1



