---
title: "ChiSqInd2.Rmd"
author: "Matthew Bardoe & Carey Kopeikin"
date: "3/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, error=TRUE, fig.width=5.3, fig.height=5 )
```



## Displaying data for Chi Square Categorical Data.
(*Note:* To make the displays look right you need to run the code at the top.)

Let's take a look at some of the data that we have looked at in the previous lesson. 

```{r}
pol <- read.csv("POL2013rev4.csv", stringsAsFactors = TRUE)

table.race.dayboarding <-table(pol$Race, pol$DayBoarding)
table.race.dayboarding

```


Lets rename the values to clean up the display. Then run the table again
```{r}
levels(pol$Race) <- c("Asian", "Black", "Other", "White") 

#Run the table again.
table.race.dayboarding <-table(pol$Race, pol$DayBoarding)
table.race.dayboarding

```


What if we wanted to combine White and Other? 
```{r}

levels(pol$Race) <- c("Asian", "Black", "Other", "Other") 

#Run the table again.
table.race.dayboarding <-table(pol$Race, pol$DayBoarding)
table.race.dayboarding

```


What if we wanted to change the order to Black, Asian, Other? 
```{r}

pol$Race <- factor(pol$Race,
                   levels = c("Black", "Asian", "Other"))

#Run the table again.
table.race.dayboarding <-table(pol$Race, pol$DayBoarding)
table.race.dayboarding

```


Here is a new plot called a Mosaic plot that can help us look at the data.

```{r}

mosaicplot(table.race.dayboarding, 
           main="Choate Day and Boarding Breakdown",
           xlab="Race",
           ylab="Day/Boarding",
           color= 1:8,
           border="yellow"
           )
```

This gives a good display that helps us see in terms of area the weight of each cell in the table. If the each row and column are broken up in a similar ways then we get the sense that the variables are independent. Specifically we are looking to see if the thickness of each colored row is the same throughout. Clearly that is not the case here. As we saw in the last class these variables are associated. 

As a reminder, when we run the chi square test on the data we see that the p-value is low enough that we can reject the null hypothesis that race and boarding status are independent and accept the alternative that they are associated.
```{r}
chisq.test(table.race.dayboarding)
```


Let's look at two variables that are independent. Before we look at the graph we will need to clean them up a bit by changing their names and changing their order. 

```{r}
table.dayBoarding.form <- table(pol$DayBoarding, pol$GRADENOPG)
table.dayBoarding.form

```


*Change the names of the grades so that they are the first two letters of each word (Freshman should become Fr). Note that you can not change the name and the order in the same step.*

```{r}

```



*Change the order of the grades so that they are in ascending order*
```{r}



```    


*Combine Fr and So to make Underclass*

```{r}


```



Now we can make a mosaic chart to visualize this data.

```{r}

           
```


We see that these aren't exactly balanced, but there is a lot of consistency. Let's see how this compares to the chi square results.


```{r}


```

Now we see that the p-value is very high and so we cannot reject the null hypothesis that the values are independent.



We can also create Mosaic plots that visually represent which sections are unusual. To do this we use the ```shade``` variable in ```mosaicplot```. Set the value of the that variable to be ```TRUE``` or put in a vector of at most 5 positive numbers to give the cut points for the shading.

Here we will look at the shading of the plot comparing Day/Boarding by Form.

```{r}
mosaicplot(table.dayBoarding.form, 
           main="Day/Boarding by Form",
           xlab="Day/Boarding",
           ylab="Grade",
           shade= c(1,2,3,4,5),
           border="black")
```

Every cell is white which means that none are weird. Being not weird is exactly what we expect with such a large p-value. 

Let's look at how a shaded mosaic plot looks like in the first graph we looked at where the variables were associated. 

```{r}
mosaicplot(table.race.dayboarding,
           main="Choate Day and Boarding Breakdown",
           xlab="Race",
           ylab="Day/Boarding",
           shade= c(1,2,3,4,5),
           border="black")

chisq.test(table.race.dayboarding)

```

In this graph the red areas are areas that are smaller than would be expected from independence and the blue are larger.

We needed to tell the cuts for the shading here because the number of categories means that the chi square value won't be too big, and therefore the residuals are smaller than most chi square tests.

*Your turn*

You will be looking at an example data set on financial behaviors and financial literacy from The National Financial Capability Study. The study is performed by the FINRA Investor Education Foundation. The survey was administered in 2009, 2012, and 2015. The 2015 study includes responses by more than 27,000 individuals from every demographic group across the United States.

The data set ```financialdifficulty.csv``` below includes a subset of the variables and observations from the full study. This subset includes 98 financial and demographic variables on 250 randomly selected individuals across the country.

The code below downloads the data and opens it, creating a data frame object, df.
```{r}
fdiff <- read.csv("financialdifficulty.csv", stringsAsFactors = TRUE)

head(fdiff)
```

Your assignment is to create displays (mosaicplots) that show how variables, ```SpendingIncome```(which asks people if they spend more, less than, or equal to their income), ```Age```, and ```ComeUp2K``` (which asks people if they can come up with $2000 in an emergency),  connect to the variable ```FinDifficulty```. Create plots, calculate and chi square tests and then write a description of the strength of the association between ```FinDifficulty``` and the given variable. Make sure your plots look good and the factor levels are in the proper orders.

Because the small number of people that say 'Very Difficult' to Financial Difficulty you may need to combine the 'Very Difficult' and 'Somewhat Difficult' categories. This will help the warning on the Chi Square test go away. 


*Create table for Spending Income vs FinDifficulty*

```{r}


```


*Create a colored mosaic plot for Spending Income vs FinDifficulty *


```{r}


```
*Can you easily read the chart? If not change the names of the levels that are too long (like we did at the very top of these guided notes) and rerun the code for the table and the mosaic plot*


```{r}


```


*After looking at the chart above do the variables Spending Income and Financial Difficulty appear to be independent or associated? Why?*





*Create a shaded mosaic plot for Spending Income vs FinDifficulty *
```{r}


```
*After looking at the chart above do the variables Spending Income and Financial Difficulty appear to be independent or associated? What areas are different than expected if the variables were independent?*





*Write down the null and alternative hypotheses for a chi square test of independence for Spending Income and Financial Difficulty*




*Run a chi square test for Spending Income and Financial Difficulty *
```{r}


```

*Interpret the results of the chi square test*






#Age vs FinDifficulty


*Create table for Age vs FinDifficulty*

```{r}


```



*Create a colored mosaic plot for Age vs FinDifficulty *


```{r}


```
*Can you easily read the chart? If not change the names of the levels that are too long (like we did at the very top of these guided notes) and rerun the code for the table and the mosaic plot*



```{r}


```



*After looking at the chart above do the variables Age and Financial Difficulty appear to be independent or associated? Why?*





*Create a colored mosaic plot for Age and Financial Difficult *
```{r}


```
*After looking at the chart above do the variables Age and Financial Difficult appear to be independent or associated? Why?*





*Write down the null and alternative hypotheses for a chi square test of independence for Age and Financial Difficulty*




*Run a chi square test for Age and Financial Difficulty *
```{r}


```
*If there is a warning that the approximation may be incorrect check the expected values. Each cell in that table must be greater than or equal to 5.* 


```{r}


```
*If they are not greater than or equal to 5 you have to combine some categories. In this case combine Somewhat difficult and Very difficult.*


```{r}

```

*Now go back to the start of the Age vs FinDifficulty section and cut and paste each code chunk into the chunk below*

```{r}


```


*Interpret the results of the chi square test*





#FinDifficulty vs ComeUp2K


*Create table for ComeUp2K vs FinDifficulty*

```{r}


```






*Create a colored mosaic plot for ComeUp2K vs FinDifficulty *


```{r}


```
*Can you easily read the chart? If not change the names and/or order of the levels that are too long (like we did at the very top of these guided notes) and rerun the code for the table and the mosaic plot*



```{r}


```



*After looking at the chart above do the variables ComeUp2K and Financial Difficulty appear to be independent or associated? Why?*





*Create a colored mosaic plot for Financial Difficulty vs ComeUp2K *
```{r}

```

*After looking at the chart above do the variables Financial Difficulty and ComeUp2K appear to be independent or associated? Why?*





*Write down the null and alternative hypotheses for a chi square test of independence for ComeUp2K and Financial Difficulty*




*Run a chi square test for ComeUp2K and Financial Difficulty *
```{r}


```


*Which areas are the most unusual?*


```{r}


```


*Interpret the results of the chi square test*













