---
title: "Day 22 Introduction to ANOVA"
author: "Matthew Bardoe & Carey Kopeikin"
date: "2/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# ANOVA: Analysis of Variance

This technique is most useful when you want to compare some quantitative data across some categorical data (often we have more than two levels to our categorical variable). We want to see if the averages were different across the different groups. Therefore we are analyzing how the mean different the internal group means are from the entire group's mean. In this way we are analyzing the variance of the means. 

A few years ago a stats class did a survey that asked 249 Choate students about their stress levels. To do this we used the DASS protocol. DASS stands for Depression, Anxiety, Stress Scale and from those questions were able to create a score for each student that took our survey. 

For instance in our Pace of Life survey. We could look at the DassTotal when we look at different grade level. The question we are answering here is if there is a significant difference between the means across these groups. 

So if we spoke about this in the language of hypothesis testing our hypothesis would be:

    * H_0: All the means of the different groups are equal
    * H_a: At least one pair of means of the groups are different
    
There are some assumptions that the data must meet to make the test be valid. These are:

* Simple Random Sampling or independent observations
* Independent groups
* Standard deviation across groups are the same
* Each group has a large enough number of observations (n>20). This supports normality.

## Pace of life example



Let's load the data and look at the information:

```{r}

pol <- read.csv("POLFINALRevise2019V2.csv")
dim(pol)
head( pol[ ,c("DassTotal","GRADENOPG")] )
```


Next we should look at the data. Below you will see a boxplot of the data grouped by various grade levels. 

```{r}
# Box plot
boxplot( DassTotal ~ GRADENOPG, data = pol,
        xlab = "Grade",
        ylab = "Stress",
        frame = FALSE,
        col = c("lightblue2", "gold", "pink", "green"))

```
The data is arranged alphabetically but we want it increasing by grade level. We can use the code we learned in the factors lab:

```{r}
# Create vector with the correct ordered levels

correct.order.grade <- c( "Freshman", "Sophomore", "Junior", "Senior" ) 


# Create a factor using the character vector with the original values
# and the ordered level vector created in step 1

pol$GRADENOPG <- factor( pol$GRADENOPG,
                         levels = correct.order.grade,
                         ordered = TRUE)

```



We can see if this worked by remaking the boxplot:

```{r}
# Box plot
boxplot( DassTotal ~ GRADENOPG, data = pol,
        xlab = "Grade",
        ylab = "Stress",
        frame = FALSE,
        col = c("lightblue2", "gold", "pink", "green"))

```
Let's make sure that we have the right kind of data for an ANOVA:

```{r}
class(pol$DassTotal)
class(pol$GRADENOPG)
levels(pol$GRADENOPG)
```


### Assumptions

The assumptions for the ANOVA say that we think that variation between groups is the same. The test is fairly robust on this point, but we should look at the standard deviations within groups just to be sure.

```{r}

tapply(pol$DassTotal, FUN=sd, INDEX=pol$GRADENOPG)
```

So there is a pretty big difference in the DASS scores variation between freshman and other grades, but let's push on.

Let's also just record what the different means are (use tapply again, but alter it a little to find the means):

```{r}
tapply(pol$DassTotal, FUN=mean, INDEX=pol$GRADENOPG)
tapply(pol$DassTotal, FUN=length, INDEX=pol$GRADENOPG)

```

#### Computing ANOVA

The command ```aov``` computes much of what we need for the ANOVA. You need to tell it what we are trying to predict, and what goes into that prediction. In this case the thing we will trying to predict will be a quantitative variable and the predictor will be categorical.


Now we do the ANOVA and save it in a variable called anov1:

```{r}
anov1 <- aov(pol$DassTotal~pol$GRADENOPG)
anov1
summary(anov1)
```

The main point of the ANOVA test is going to be the p-value which is reported in the last column of the table. Here it is 0.0304. This is less than 0.05 value that we often set as a cutoff for whether we would reject our null hypothesis that the average level of stress among grades is the same. 


#### What does it all mean?

So we see from ANOVA that the means are different. That isn't very actionable information. Are they all different from each other, or is there just one class that feels more stress. ANOVA doesn't do that kind of analysis. But there are other tests that do, and once we have done ANOVA, and seen something there we should dig deeper. 

One way to do this is the Tukey Honest Significant Differences,  ```TukeyHSD())```.

```{r}
TukeyHSD(anov1)
```

This complex table tells us about all the comparisons from different grades. There are 6 of them. We first scan over the p-adj values looking for values that are less than 0.05. These are the most significant differences. What this table is telling us, is that there does seem to be a difference between the stress levels of freshman and sophomores. The diff value is 2.30, indicating that the DASS Score is expected to go up by 2.3 for sophomores versus Freshman. But that is on the only definitively low p-value. The only other close one is the comparison of Junior and Freshman. This jibes well with our observations from the boxplot at the start. 

### Your Turn

The dataset ```chocolate.csv``` has data about different chocolate bars. In this dataset there three different kinds of chocolate beans used in the production of the chocolate bars tested: Trinitario, Criollo, and Forastero. Each chocolate bar is also given a rating. We are interested if the mean rating of the chocolate bar depends on the type of bean used.

1. Make a boxplot investigating the ratings to the type of bean.



2. Calculate the mean rating for each type of bean and the standard deviation ratings by bean.



3. Compute the ANOVA test for the mean rating by bean type. Discuss the meaning of the p-value.


4. Create a plot for the residuals and see if there are outliers.


5. Create a plot testing the normality discussion.


6. Look for pairwise differences in the rating levels by bean type.

